Almost compatible functions and infinite length games

### 2015-11-24 articles/almost-compatible-functions-and-topological-games/article.tex
fleshed out topology construction proof

@@

  \begin{proposition}
    Let \(X\) be a \(T_2\) locally countable \(\omega\)-bounded space.
    Then for each \(A\subseteq X\), \(\cl A\) has the same cardinality and
    weight as \(A\).
  \end{proposition}

  %TODO?

  \begin{lemma}
    Let \(X\) be a \(T_2\) space of cardinality less than
    \(\aleph_\omega\) which is locally countable, locally compact,
    and the closure of each set has the same cardinality as the set.
    Then \(X\) has an \(\omega\)-bounded extension \(\tilde X\)
    with the same properties,
    such that \(\tilde X\setminus X\) has the same cardinality
    as \(X\).
  \end{lemma}

  \begin{proof}
    We prove this by induction on \(n\).
    If \(n=0\), then we can just use the one-point compactification of two
    copies of \(X\). So suppose \(n>0\) and
    that \(X=\omega_n\) has such a topology. Note that
    \(X\) has a base of compact open neighborhoods
    with countable cardinality. To see this,
    for each point \(x\) let \(K\) be a countable and compact neighborhood
    of \(x\), and it follows that it is contained in a countable, open,
    and locally compact neighborhood \(W\) of \(x\), which in turn
    is zero-dimensional and metrizable. So choose \(V\) clopen
    in \(W\) such that \(x\in V\subseteq K\); \(V\) is a compact open
    neighborhood of \(x\) in \(X\).

    For each \(\alpha<\omega_n\), \(\gamma_\alpha\) may be chosen such that
    both the closure of the set \(\alpha\) in \(X\) and a countable
    neighborhood of the point \(\alpha\) are subsets of \(\gamma_\alpha\).
    Note that the set
    \(\{\lambda<\omega_n:\alpha<\lambda\Rightarrow\gamma_\alpha<\lambda\}\)
    is a cub subset of \(\omega_n\) with a cub subset
    \(C\) containing only limits of limit ordinals.
    Now for each \(\lambda\in C\), the initial segment \(\lambda\) is
    open as \(\alpha<\lambda\) belongs to the neighborhood
    \(\gamma_\alpha\subsetneq\lambda\). Also, if \(\lambda\) has uncountable
    cofinality, then for \(\beta\geq\lambda\) and any countable neighborhood
    \(U\) of \(\beta\), \(U\cap\lambda=U\cap\alpha\)
    for some \(\alpha<\lambda\); thus \(U\setminus\cl\alpha=U\setminus\lambda\)
    is a neighborhood of \(\beta\), showing that \(\lambda\)
    is clopen.

    Let \(\tilde X=\omega_n\times 2\).
    By induction on \(\lambda \in C\) we will define a topology
    for \(\tilde X_\lambda=\omega_n\times\{0\}\cup\lambda\times\{1\}\)
    such that
    \begin{itemize}
      \item \(\omega_n \times \{0\}\) is an open copy of \(X\),
      \item \(\lambda\times 2\) is open, and also closed when \(\cf\lambda>\omega\),
      \item the space has a base of compact open neighborhoods with
        countable cardinality, and
      \item for each \(\alpha<\lambda\), the closure of \(\alpha\times 2\) is
        an \(\omega\)-bounded  subset of \(\lambda\times 2\).
    \end{itemize}
    and note that \(\omega_n\times\{0\}\cup 0\times\{1\}\) satisfies these
    requirements.

    We first consider the case \(n=1\).
    If \(\lambda\) is a limit in \(C\), then
    \(
      \omega_n\times\{0\}\cup\lambda\times\{1\}
        =
      \bigcup_{\mu\in C\cap\lambda}\omega_n\times\{0\}\cup\mu\times\{1\}
    \)
    will have the topology induced by that union which
    satisfies the induction requirements.
    Otherwise we choose an increasing sequence of limit
    ordinals \(\{\alpha_k : k\in \omega\}\) with limit \(\lambda\)
    such that \(\alpha_0\) is the predecessor of \(\lambda\) in \(C\),
    or \(\alpha_0=0\) if \(\lambda\) is the least element of \(C\).

@@

    We need only describe a base for the points
    \(\<\alpha,1\>\in(\lambda\setminus\alpha_0)\times\{1\}\);
    we do so by letting
    \(\<\alpha,1\>\) be isolated when \(\alpha\not\in\{\alpha_k:k<\omega\}\),
    and letting \(\<\alpha_k,1\>\) be a compactifying point for
    \(U_k\cup(\alpha_k\times\{1\})\).

    The first two requirements of our inductive hypothesis are obviously
    satsified. Note \(U_k\cup((\alpha_k+1)\times\{1\})\) is
    a compact open neighborhood of countable cardinality
    for points in \(\lambda\times 2\), and for points in
    \((\omega_n\setminus\cl\lambda)\times\{0\}\) we may use a compact
    open neighborhood of countable cardinality from \(X\setminus\lambda\).
    Finally let \(\<\alpha,0\>\in(\cl\lambda\setminus\lambda)\times\{0\}\)
    and let \(K\subseteq\cl\lambda\times\{0\}\cup\alpha_0\times\{1\}\) be a
    compact open neighborhood in that subspace missing \(\alpha_0\times 2\).
    Then for each \(k<\omega\), \(\<\alpha_k,1\>\) has a neighborhood missing
    the compact set \(K\cap U_k\), so \(K\) gains no new limit points in
    \(\cl\lambda\times\{0\}\cup\lambda\times\{1\}\) and thus remains compact.
    So the third requirement is also satisfied.
    For the final requirement, note that for \(\alpha<\lambda\), we may
    choose \(\alpha<\alpha_k<\lambda\) and note that
    \(\alpha\times 2\) is contained in the compact
    set \(U_k\cup((\alpha_k+1)\times\{1\})\).

@@

    As a result, \(\tilde X=\bigcup_{\lambda\in C}\tilde X_\lambda\)
    is easily seen to be locally compact and locally countable. It is also
    \(\omega\)-bounded as any countable set is contained in some
    \(\alpha\times 2\) for \(\alpha<\lambda\in C\). Finally by the previous
    proposition, the closure operation cannot increase cardinalities.
  \end{proof}

  \begin{theorem}
    For each $n\in \omega$, there is a \(T_2\) locally countable,
    $\omega$-bounded topology on $\omega_n$.
  \end{theorem}

>>>>

  \begin{proposition}
    Let \(X\) be a \(T_2\) space with a base of countable and compact
    neighborhoods.
    Then \(X\) is locally metrizable, \(X\)
    has a base of compact open countable sets,
    and its closure operation preserves cardinality and weight.
  \end{proposition}

  \begin{proof}
    For each point \(x\) let \(K\) be a countable and compact neighborhood
    of \(x\), and it follows that it is contained in a countable, open,
    and locally compact neighborhood \(W\) of \(x\), which in turn
    is zero-dimensional and metrizable. So choose \(V\) clopen
    in \(W\) such that \(x\in V\subseteq K\); \(V\) is a compact open
    neighborhood of \(x\) in \(X\).

    Since the space is locally metrizable and thus first-countable,
    the weight and cardinality of any
    subspace coincide. Then for infinite \(A\subseteq X\), cover \(\cl A\) with
    open sets of countable cardinality, and say
    \(x\sim y\) if there's a finite linked chain joining them; since
    \(x\sim a\) for some \(a\in A\), \(\cl A\) is covered by
    \(|A|\)-many sets each of size \(\aleph_0\). Thus
    \(|A|\leq|\cl A|\leq|A|\otimes\aleph_0=|A|\).
  \end{proof}

@@

  \begin{lemma}
    Let \(X\) be a \(T_2\) space of cardinality less than
    \(\aleph_\omega\) with a base of countable and compact neighborhoods.
    Then \(X\) has an \(\omega\)-bounded extension \(\tilde X\)
    with the same properties,
    such that \(\tilde X\setminus X\) has the same cardinality
    as \(X\).
  \end{lemma}

  \begin{proof}
    We prove this by induction on \(n\).
    If \(n=0\), then we can just use the one-point compactification of two
    copies of \(X\). So suppose \(n>0\) and
    that \(X=\omega_n\) has an appropriate topology.

    For each \(\alpha<\omega_n\), \(\gamma_\alpha\) may be chosen such that
    both the closure of the set \(\alpha\) in \(X\) and a countable
    neighborhood of the point \(\alpha\) are subsets of \(\gamma_\alpha\).
    Note that the set
    \(\{\lambda<\omega_n:\alpha<\lambda\Rightarrow\gamma_\alpha<\lambda\}\)
    is a cub subset of \(\omega_n\) containing a cub subset \(C\)
    of limit ordinals.
    Now for each \(\lambda\in C\), the initial segment \(\lambda\) is
    open as \(\alpha<\lambda\) belongs to the neighborhood
    \(\gamma_\alpha\subsetneq\lambda\). Also, if \(\lambda\) has uncountable
    cofinality, then for \(\beta\geq\lambda\) and any countable neighborhood
    \(U\) of \(\beta\), \(U\cap\lambda=U\cap\alpha\)
    for some \(\alpha<\lambda\); thus \(U\setminus\cl\alpha=U\setminus\lambda\)
    is a neighborhood of \(\beta\), showing that \(\lambda\)
    is clopen.

    Let \(\tilde X=\omega_n\times 2\).
    By induction on \(\lambda \in C\) we will define a topology
    for \(\tilde X_\lambda=\omega_n\times\{0\}\cup\lambda\times\{1\}\)
    such that
    \begin{itemize}
      \item \(\omega_n \times \{0\}\) is an open copy of \(X\),
      \item \(\lambda\times 2\) is open, and when \(\cf\lambda>\omega\) also closed,
      \item the space has a base of countable and compact neighborhoods, and
      \item for each \(\alpha<\lambda\), the closure of \(\alpha\times 2\) is
        an \(\omega\)-bounded  subset of \(\lambda\times 2\).
    \end{itemize}
    and note that \(\omega_n\times\{0\}\cup 0\times\{1\}\) satisfies these
    requirements.

    We first consider the case \(n=1\).
    If \(\lambda\) is a limit in \(C\), then
    \(
      \omega_n\times\{0\}\cup\lambda\times\{1\}
        =
      \bigcup_{\mu\in C\cap\lambda}\omega_n\times\{0\}\cup\mu\times\{1\}
    \)
    will have the topology induced by that union which
    satisfies the induction requirements.
    Otherwise we choose an increasing sequence of
    ordinals \(\{\alpha_k : k\in \omega\}\) with limit \(\lambda\)
    such that \(\alpha_0\) is the predecessor of \(\lambda\) in \(C\),
    or \(\alpha_0=0\) if \(\lambda\) is the least element of \(C\).

@@

    We need only describe a base for the points
    \(\<\alpha,1\>\in(\lambda\setminus\alpha_0)\times\{1\}\).
    We do so by letting
    \(\<\alpha,1\>\) be isolated when \(\alpha\not\in\{\alpha_k:k<\omega\}\),
    and giving \(\<\alpha_k,1\>\) the neighborhoods
    \(U_k\cup((\alpha_k+1)\times\{1\})\setminus K\)
    for each compact \(K\) in \(U_k\cup(\alpha_k\times\{1\})\); that is,
    let \(\<\alpha_k,1\>\) be the one point compactifying
    \(U_k\cup(\alpha_k\times\{1\})\).

    The first two requirements of our inductive hypothesis are obviously
    satsified. Note \(U_k\cup((\alpha_k+1)\times\{1\})\) is
    a compact neighborhood of countable cardinality
    for points in \(\lambda\times 2\), and for points in
    \((\omega_n\setminus\cl\lambda)\times\{0\}\) we may use a compact
    neighborhood of countable cardinality from \(X\setminus\lambda\).
    Finally let \(\<\alpha,0\>\in(\cl\lambda\setminus\lambda)\times\{0\}\)
    and let \(K\subseteq\cl\lambda\times\{0\}\cup\alpha_0\times\{1\}\) be a
    compact neighborhood in that subspace missing \(\alpha_0\times 2\).
    Then for each \(k<\omega\), \(\<\alpha_k,1\>\) has the neighborhood
    \(U_k\cup((\alpha_k+1)\times\{1\})\setminus K\),
    so \(K\) gains no new limit points in
    \(\cl\lambda\times\{0\}\cup\lambda\times\{1\}\) and thus remains compact.
    So the third requirement is also satisfied.
    For the final requirement, note that for \(\alpha<\lambda\), we may
    choose \(\alpha<\alpha_k<\lambda\) and note that
    \(\alpha\times 2\) is contained in the compact
    set \(U_k\cup((\alpha_k+1)\times\{1\})\).

@@

    As a result, \(\tilde X=\bigcup_{\lambda\in C}\tilde X_\lambda\)
    is
    \(\omega\)-bounded as any countable set is contained in some
    \(\alpha\times 2\) for \(\alpha<\lambda\in C\).
  \end{proof}

  \begin{theorem}
    For each $n\in \omega$, there is a \(T_2\), locally countable,
    $\omega$-bounded topology on $\omega_n$.
  \end{theorem}
